In mammals, 20 to 40\% of advantageous mutations are restoring damaged genomes instead of creating adaptive innovations

### 2022-07-12 manuscript/main.tex
Refactored Supp. Mat. and included explanations

    We replicated such experiment across 28 populations (fig.~\ref{fig:heatmap}), and found that mutations predicted to be deleterious are indeed purged at the population-genetic scale (blue and green rows in fig.~\ref{fig:heatmap}A).
    However, mutations predicted to be weakly advantageous ($\divWeakAdv$) are composed of a mix of neutral and selected mutations with varying proportions across the different populations (orange row in fig.~\ref{fig:heatmap}B).
    Similarly, mutations predicted to be advantageous are found to be positively selected with varying proportions across populations (red row in fig.~\ref{fig:heatmap}C).
    These variable proportions between populations can be explained by the effective number of individuals in the population ($\Ne$), a major driver of selection efficacy.
    Since $\Ne$ is not directly measurable, a reasonable proxy usually used is the synonymous diversity (see material and methods).
    We thus ordered populations by their levels of synonymous diversity (fig.~\ref{fig:diversity}A) and confirmed that higher synonymous diversity is typically accompanied by a smaller proportion of neutral mutations.
    This result suggests that populations with higher diversity (e.g.~\textit{Bos} or \textit{Ovis}) are more likely to discriminate whether mutations are advantageous or deleterious (fig~S1-S5).
    Alternatively stated, mutations in populations with low diversity (e.g.~\textit{Homo}) are effectively nearly-neutral and behave as would a neutral mutation.
    %, especially for mutations already predicted to be nearly-neutral at the phylogenetic scale (fig~S8-S9).
    This result is qualitatively in accordance with the nearly-neutral theory of evolution which argues that mutations are less efficiently selected for in small populations.

@@

    %Altogether, we estimated that between 18 and 39\% of $\dnds$ is over-estimated (table S1), corresponding to maintenance positive selection.
    %In a nearly neutral framework, genomes could be viewed as the arena of a tug-of-war between a large volume of deleterious mutations that eventually reach fixation due to drift, and scarce advantageous mutations that are restoring a more optimal ancestral state.
    % However, and surprisingly, a decrease in the proportion of neutral mutation did not translate into a mechanical increase in the proportion of advantageous mutations.
    % Instead, populations with higher diversity have a lower proportion of advantageous mutations (fig.~S7), this effect is also more pronounced for mutations predicted to be advantageous at the phylogenetic scale (red row in fig.~\ref{fig:heatmap}C and fig.~S8).
    % This discrepancy can be explained by the fact that populations with larger diversity have already depleted the putatively advantageous mutations, and that only deleterious are remaining possible.
    % In other words, being on the top of a fixed fitness landscape leaves you with only paths going downhill.
    % Altogether, populations with increased diversity are showing signature of increased selection efficiency and are also less likely to generate positively selected mutations.

@@

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.heatmap}
        \caption{
            Reproducibility of results shown in figure~\ref{fig:homo-afr-results} in 28 populations across 6 genera.
            Site-frequency spectra and total mutation rates are combined at the population-genetic scale to estimate the proportion of advantageous mutations ($\PpolyAdv$) in the top panel, of nearly-neutral mutations ($\PpolyNeutral$) in the middle panel and of deleterious mutations ($\PpolyDel$) in the bottom panel.
        }
        \label{fig:heatmap}
    \end{figure*}

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.diversity}
        \caption{
            Panel A. For each population, the proportion of advantageous ($p_+=\PpolyAdv$), nearly-neutral ($\PpolyNeutral$) and of deleterious ($\PpolyDel$) mutations estimated at the population-genetic scale is shown.
            Estimation is performed with SNPs predicted to be weakly advantageous at the phylogenetic scale ($\divWeakAdv$).
            Panel B. $R(p_+)=\dfrac{p_+ - p_+(\Sphy < 0)}{p_+}$ is the fraction of non-adaptive advantageous mutations, $p_+$ is compared to its estimation when we had removed all the restoring mutations $p_+(\Sphy < 0)$.
            $R(p_+)$ (y-axis) is shown as a function of the synonymous diversity (Watterson $\theta_w$ in x-axis).
            Panel C. $R(\dnds)=\dfrac{\dnds - \dn(\Sphy < 0) / \ds}{\dnds}$ is the fraction of divergence ($\dnds$) that is due to maintenance positive selection, where $\dnds$ is compared to the estimated divergence when we removed the restoring substitutions $\dn(\Sphy < 0)/\ds$.
            $R(\dn)$ (y-axis) is shown as a function of the synonymous diversity (Watterson $\theta_w$ in x-axis).
        }
        \label{fig:diversity}
    \end{figure*}

    % Our study is a bridge between phylogenetic and population genetics
    From a theoretical evolution perspective, we have found that we can use phylogenetic signals and mutation-selection codon models to predict the selection coefficient of a mutation.
    However, we so far assumed no changes in effective population size while it has been established that population size changes have a major effect on selection dynamics\cite{lanfear_population_2014, jones_shifting_2017, platt_protein_2018}.
    Estimating effective population sizes along with fitness landscapes is possible\cite{latrille_inferring_2021}, but is, for the moment, too computational intensive to be performed genome-wide\cite{latrille_inferring_2021}.
    Hybrid mutation-selection and $\omega$-based codon models could provide a way forward\cite{brevet_reconstructing_2021}.
    Moreover, SIFT scores based on amino-acid alignments across species have also been found to be informative of the selection pressure exerted at the population scale\cite{chen_hunting_2021}.
    Accordingly, our study shows that sites with a higher SIFT score at the phylogenetic scale are more positively selected in the different populations (supp.\ mat\  section 2).
    In practice, SIFT scores are more coarse grained, with for example a bulk of mutations with a score of 1.0 (the highest), and does not provide a boundary between advantageous and deleterious mutations, nor are they based on a population-genetics formalism.

@@

    At the population-genetic level, our study relied on models with some shortcomings, such that there is no consensus on the expected shape of the DFE\cite{welch_divergence_2008, bataillon_effects_2014}, although it has been found to be reliably constant between species\cite{castellano_comparison_2019}.
    The model we used allows the inference of a full DFE, modelling the negative part of the DFE as a gamma distribution, and the positive part as an exponential distribution.
    However, when we aggregate SNPs that are expected to be strongly deleterious or strongly advantageous, the model struggles to fit their DFE to a gamma or an exponential distribution.\cite{brevet_reconstructing_2021}
    As a result, while the proportions of negative/weak/positive mutations stays reliable, the mean values for selection coefficients ($\Spop$) estimated by polyDFE become biologically absurd ($\Spop \ll -10^3$).
    Such effect is not found for neutral SNPs and a method to accurately estimate the mean value of selection coefficients from SFS without assuming any shape for the DFE remains yet to be developed.
    Aside from methodological limitations, the mammalian case study might also not be representative of other clades.
    Indeed, mammals have relatively small long term population sizes, which allow large amounts of deleterious mutations to be fixed, and thus, there are a lot of opportunities for maintenance positive selection to arise.
    It would thus be of great interest to reproduce our experiment with other clades, such as \textit{Drosophila} or birds which tend to have higher effective population size.
    It would also be interesting to have access to more shallow but wider phylogenies to increase the resolution of our predictions.
    % Moreover, mammalian are less affected by codon usage

>>>>

    We replicated such experiment across 28 populations (fig.~\ref{fig:heatmap}), and found that mutations predicted to be deleterious are indeed purged at the population-genetic scale (blue and green rows in fig.~\ref{fig:heatmap}A).
    However, mutations predicted to be weakly advantageous ($\divWeakAdv$) are composed of a mix of neutral and selected mutations with varying proportions across the different populations (orange row in fig.~\ref{fig:heatmap}B).
    Similarly, mutations predicted to be advantageous are found to be positively selected with varying proportions across populations (red row in fig.~\ref{fig:heatmap}C).
    These variable proportions between populations can be explained by the effective number of individuals in the population ($\Ne$), a major driver of selection efficacy.
    Since $\Ne$ is not directly measurable, a reasonable proxy usually used is the synonymous diversity (see material and methods).
    We thus ordered populations by their levels of synonymous diversity (fig.~\ref{fig:diversity}A) and confirmed that higher synonymous diversity is typically accompanied by a smaller proportion of neutral mutations.
    This result suggests that populations with higher diversity (e.g.~\textit{Bos} or \textit{Ovis}) are more likely to discriminate whether mutations are advantageous or deleterious (fig~S1-S5).
    Alternatively stated, mutations in populations with low diversity (e.g.~\textit{Homo}) are effectively nearly-neutral and behave as would a neutral mutation, especially for mutations already predicted to be nearly-neutral at the phylogenetic scale (fig~S8-S9).
    This result is qualitatively in accordance with the nearly-neutral theory of evolution which argues that mutations are less efficiently selected for in small populations.

@@

@@

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.heatmap}
        \caption{
            Reproducibility of results shown in figure~\ref{fig:homo-afr-results} in 28 populations across 6 genera.
            For each class of selection at the phylogenetic scale ($\Sphyclass \in \{ \divStrongDel, \divDel,  \divWeakDel,  \divWeakAdv, \divAdv \}$, in rows), site-frequency spectra and total number of sites are combined at the population-genetic scale to estimate the proportion of advantageous mutations ($\PpolyAdv$) in the top panel, of nearly-neutral mutations ($\PpolyNeutral$) in the middle panel and of deleterious mutations ($\PpolyDel$) in the bottom panel.
        }.
        \label{fig:heatmap}
    \end{figure*}

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.diversity}
        \caption{
            Panel A. For each population, the proportion of advantageous ($p_+=\PpolyAdv$), nearly-neutral ($\PpolyNeutral$) and of deleterious ($\PpolyDel$) mutations estimated at the population-genetic scale using all SNPs available.
            Population are sorted for increased synonymous diversity.
            Panel B. $R(p_+)=\dfrac{p_+ - p_+(\Sphy < 0)}{p_+}$ is the fraction of non-adaptive advantageous mutations, $p_+$ is compared to its estimation when we had removed all the restoring mutations $p_+(\Sphy < 0)$.
            $R(p_+)$ (y-axis) is shown as a function of the synonymous diversity (Watterson $\theta_w$ in x-axis).
            Panel C. $R(\dnds)=\dfrac{\dnds - \dn(\Sphy < 0) / \ds}{\dnds}$ is the fraction of divergence ($\dnds$) that is due to maintenance positive selection, where $\dnds$ is compared to the estimated divergence when we removed the restoring substitutions $\dn(\Sphy < 0)/\ds$.
            $R(\dn)$ (y-axis) is shown as a function of the synonymous diversity (Watterson $\theta_w$ in x-axis).
        }
        \label{fig:diversity}
    \end{figure*}

    % Our study is a bridge between phylogenetic and population genetics
    From a theoretical evolution perspective, we have found that we can use phylogenetic signals and mutation-selection codon models to predict the selection coefficient of a mutation.
    However, we so far assumed no changes in effective population size while it has been established that population size changes have a major effect on selection dynamics\cite{lanfear_population_2014, jones_shifting_2017, platt_protein_2018}.
    Estimating effective population sizes along with fitness landscapes is possible\cite{latrille_inferring_2021}, but is, for the moment, too computational intensive to be performed genome-wide\cite{latrille_inferring_2021}.
    Hybrid mutation-selection and $\omega$-based codon models could provide a way forward\cite{brevet_reconstructing_2021}.
    Moreover, SIFT scores based on amino-acid alignments across species have also been found to be informative of the selection pressure exerted at the population scale\cite{chen_hunting_2021}.
    Accordingly, our study shows that sites with a higher SIFT score at the phylogenetic scale are more positively selected in the different populations (supp.\ mat\  section SIFT scores).
    In practice, SIFT scores are more coarse grained, with for example a bulk of mutations with a score of 1.0 (the highest), and does not provide a boundary between advantageous and deleterious mutations, nor are they based on a population-genetics formalism.

@@

    At the population-genetic level, our study relied on models with some shortcomings, such that there is no consensus on the expected shape of the DFE\cite{welch_divergence_2008, bataillon_effects_2014}, although it has been found to be reliably constant between species\cite{castellano_comparison_2019}.
    The model we used allows the inference of a full DFE, modelling the negative part of the DFE as a gamma distribution, and the positive part as an exponential distribution.
    However, when we aggregate SNPs that are expected to be strongly deleterious or strongly advantageous, the model struggles to fit their DFE to a gamma or an exponential distribution.\cite{brevet_reconstructing_2021}
    As a result, while the proportions of negative/weak/positive mutations stays reliable, the mean values for selection coefficients ($\Spop$) estimated by polyDFE become biologically absurd ($\Spop \ll -10^3$).
    Such effect is not found for neutral SNPs and a method to accurately estimate the mean value of selection coefficients from SFS without assuming any shape for the DFE remains yet to be developed.
    Aside from methodological limitations, the mammalian case study might also not be representative of other clades.
    Indeed, mammals have relatively small long term population sizes, which allow large amounts of deleterious mutations to be fixed, and thus, there are a lot of opportunities for maintenance positive selection to arise.
    It would thus be of great interest to reproduce our experiment with other clades, such as \textit{Drosophila} or birds which tend to have higher effective population size.
    It would also be interesting to have access to more shallow but wider phylogenies to increase the resolution of our predictions.
